Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-03-31 article/04_Discussion.Rmd
and more phrasing

The second important question posed for this paper was, *how cultural distance between regions that results from changes in burial traditions develops, and how it is related to spatial distance?* While the idea behind the first question was rather technical -- reproducing already available information in a quantitative way -- the second one had a larger scope. The quantitative representation of an archaeological narrative paths the way for a diverse set of methods to be applied and the analysis undertaken here is merely the beginning.

@@

Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:dacta} and \ref{tab:dactb}, Figure \ref{correlationtimeseries}). Only for certain periods of time this correlation is visible in the Radon-B data. The *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation. If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000 calBC) and a decline (after 2000 calBC) of cultural-spatial-distance correlation. There are several reasons why any interpretation of this observation is difficult so far:

1. The amount, distribution and quality of the data as discussed in section \ref{data-quality-discussion} question the credibility of the cultural distance calculation. The macro regions have to be very big due to a lack of data and don't allow for a more precise picture. The differences in the number of available dates -- sample size -- per region introduce sampling error. The samples are also non-independent, as large sites with multiple burials can strongly bias the proportions of a particular region. Furthermore the region size also causes strong risk of Simpson’s paradox (i.e. a trend appears in several different groups of data but disappears or reverses when these groups are combined).

>>>>

The second important question posed for this paper was, *how cultural distance between regions develops, and how it is related to spatial distance?* While the idea behind the first question was rather technical -- reproducing already available information in a quantitative way -- the second one had a larger scope. The quantitative representation of an archaeological narrative paths the way for a diverse set of methods to be applied and the analysis undertaken here is merely the beginning.

@@

Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:dacta} and \ref{tab:dactb}, Figure \ref{correlationtimeseries}). Only for certain periods of time this correlation is visible in the Radon-B data and the *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation. If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000 calBC) and a decline (after 2000 calBC) of cultural-spatial-distance correlation. There are several reasons why any interpretation of this observation is difficult so far:

1. The amount, distribution and quality of the data as discussed in section \ref{data-quality-discussion} question the credibility of the cultural distance calculation. The macro regions have to be very big due to a lack of data and don't represent the cultural heterogeneity of Bronze Age Europe appropriately. The differences in the number of available dates -- sample size -- per region introduce sampling error. The samples are also non-independent, as large sites with multiple burials can strongly bias the proportions of a particular region. Furthermore the region size also causes strong risk of Simpson’s paradox (i.e. a trend appears in several different groups of data but disappears or reverses when these groups are combined).
